---
title: "GloBAM Data Management Plan"
author:
- Peter Desmet
- Cecilia Nilsson
- Adriaan Dokter
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
github-repo: "enram/globam-dmp"
output:
  bookdown::gitbook:
    split_by: rmd
    df_print: kable
    config:
      toc:
        collapse: section
      search: yes
      edit: https://github.com/enram/globam-dmp/blob/master/%s
      download: ["pdf"]
      sharing:
        facebook: no
        twitter: yes
        all: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, # Exclude code chunks from output
  warning = FALSE,
  message = FALSE
)

# Load required libraries
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(lubridate)      # To work with dates
library(here)           # To find files
library(janitor)        # To add totals to tables
library(jsonlite)       # To read json
```

<!-- References used in this document -->

[enram-repo]: http://enram.github.io/data-repository/

# Intro

Welcome to the data management plan (DMP) for the project _GloBAM: Towards monitoring, understanding and forecasting global biomass flows of aerial migrants_.

## Distributions

- Website: <https://enram.github.io/globam-dmp/>
- PDF: <https://enram.github.io/globam-dmp/globam-dmp.pdf>

## How this DMP is maintained

1. This DMP is maintained and versioned on GitHub at <https://github.com/enram/globam-dmp/>.
2. Each chapter is an [R Markdown](https://rmarkdown.rstudio.com/) file (`Rmd`) in the `src` directory of that GitHub repository. You can access it directly by clicking the pencil icon in the top navigation of this website.
3. Changes to the R Markdown files can be made by contributors to the GitHub repository or suggested by anyone as pull requests. Textual changes can be done directly on GitHub, code changes are better tested in RStudio first.
4. Accepted changes (i.e. changes to the `master` branch) will trigger an [automatic build procedure](https://travis-ci.org/enram/globam-dmp) that will generate a new version of this DMP using the R package [bookdown](https://bookdown.org/yihui/bookdown/). The date of the build is used as the version number.

## Citation

```{r}
# Read metadata from index.Rmd
frontmatter <- rmarkdown::yaml_front_matter(here("index.Rmd"))

# Split authors in form Last F, Last F
authors <- map(frontmatter$author, function(x) {
  split_name <- str_split(x, pattern = " ")[[1]]
  paste(split_name[2], str_sub(split_name[1], 1, 1))
})

# Create citation
citation <- paste0(
  paste(authors, collapse = ", "),
  " (",
  year(Sys.Date()),
  ") ",
  frontmatter$title,
  ". Version ",
  Sys.Date(),
  ". https://enram.github.io/globam-dmp/"
)
```

> `r citation`
