# (PART\*) Source data {-}

# European radar data (`pvol`)

This chapter describes polar volume (`pvol`) data from European weather radars.

## Source

`pvol` data are collected by **national weather services** in Europe. They have different approaches in processing, archiving and providing access to these data. Some **research labs** (e.g. UvA) also have archived subsets of these `pvol` data for research use.

The best source for consolidated `pvol` data however is **[BALTRAD](http://baltrad.eu/)**, managed by [SMHI](https://www.smhi.se/en/research/research-departments/atmospheric-remote-sensing/contact-us-1.4801) (contact person: Günther Haase). `pvol` data are archived there, but access to that server is restricted to SMHI. See \@ref(baltrad-pvol) for details on how `pvol` data are submitted to this archive.

The rest of this chapter discusses the specifics of the BALTRAD archive.

## License

European radar data exchange is coordinated by the _European Operational Program for Exchange of Weather Radar Information ([EUMETNET/OPERA](http://eumetnet.eu/activities/observations-programme/current-activities/opera/))_. GloBAM has access to these data via the _EIG EUMETNET license agreement for Research use of OPERA data_, which grants a specific list of people (involved in ENRAM and GloBAM) access to radar data from selected countries under the following conditions (excerpt):

> The Licensor grants the Licensees a non-Exclusive license to use intellectual property belonging to the EUMETNET members as defined below [ ... ]
> 
> The OPERA Members Data will be provided for use within the ENRAM Project for the purpose of extracting animal migration information for scientific research.
> 
> The Grant of this License does not permit use of the OPERA Members Data licensed to be used for commercial purposes or exploitation for profit.

The license agreement allows us to access European `pvol` data for the sole purpose of extracting animal migration information for scientific research. Storing, sharing or other use of these data is **restricted** and requires prior written consent of the Licensor.

## Format

The delivery of European `pvol` data is defined in the license agreement:

> The Licensor will make the OPERA Members Data available to the Licensees in accordance to the following data description and technical specifications:
> 
> a. Single-site polar volumes containing reflectivity, optionally also Doppler velocity, uncorrected reflectivity and dual-pol parameters. Data model ODIM as described in OPERA pages of the EUMETNET website (http://www.eumetnet.eu)
> b. Update frequency – 15 minutes
> c. Issue time – up to 120 minutes after data time
> d. Format: HDF5
> e. Delivery method: FTP via Internet using an ad hoc server. A username and password will be created for ENRAM.
> f. Availability of OPERA Members Data  will be on the basis of reasonable endeavour
> g. Support: OPERA documents are available from the EUMETNET website’s OPERA pages (http://www.eumetnet.eu). OPERA Program Manager will manage technical questions related to the OPERA Products but may address more complicated issues to a team member who may charge a fee.

Data is thus provided for every 15 minutes as `hdf5` in the [OPERA ODIM_h5 format](https://github.com/adokter/vol2bird/blob/master/doc/ODIM_H5_v23.pdf). These files can be read in `bioRad` with [`read_pvolfile()`](http://adokter.github.io/bioRad/reference/read_pvolfile.html).

## Geographical scope

```{r download_opera_db}
# Download OPERA database as json and store in this repo
download.file("http://eumetnet.eu/wp-content/themes/aeron-child/observations-programme/current-activities/opera/database/OPERA_Database/OPERA_RADARS_DB.json", here("data", "OPERA_RADARS_DB.json"))
```

```{r}
# Read data files
opera_radars <- fromJSON(here("data", "OPERA_RADARS_DB.json"))
opera_countries <- read_csv(here("data", "opera_countries.csv"))
```

OPERA manages a list of radar metadata (`r opera_radars %>% nrow()` radars). This **OPERA database** is available as [an interactive map](http://eumetnet.eu/wp-content/themes/aeron-child/observations-programme/current-activities/opera/database/OPERA_Database/index.html) derived from [a json file containing the metadata](http://eumetnet.eu/wp-content/themes/aeron-child/observations-programme/current-activities/opera/database/OPERA_Database/OPERA_RADARS_DB.json). To keep track of changes, we archive this `json` on [GitHub](https://github.com/enram/globam-dmp/blob/master/data/OPERA_RADARS_DB.json) every time we update the DMP.

The license agreement allows access to data from `r opera_countries %>% filter(in_license_agreement) %>% nrow()` countries:

```{r}
# Show table of countries
opera_radars %>%
  # Add country information
  left_join(opera_countries, by = "country") %>%
  
  # Filter on countries in license agreement
  filter(in_license_agreement) %>%
  
  # Group by country
  group_by(country, iso_code) %>%
  arrange(country) %>%
  
  # Show total number of radars + operational radars (status = 1)
  summarize(
    radars = n_distinct(number),
    operational = n_distinct(number[status == 1])
  ) %>%
  adorn_totals("row")
```

The BALTRAD archive does not contain `pvol` data for all of these radars/countries however (see \@ref(baltrad-pvol)). Its precise geographical scope is currently **unknown**, but being gathered in [this spreadsheet](https://docs.google.com/spreadsheets/d/1dulZ4TLjxIjmUek5QP86uJ9DsZqm9d0G6vl6hnA0mw4/edit?usp=sharing).

## Temporal scope

The temporal scope of the BALTRAD archive is currently **unknown**.

## Historical archive

GloBAM needs a **2-3 year archive** of European `pvol` and derived `vp` data to tackle its research questions. By starting from a `pvol` archive, the generation of `vp` data can be reproduced if need be, e.g. to make sure adequate `vol2bird` settings/versions are used.

The `pvol` archive at BALTRAD is likely to be the best source for this archive, but its scope and quality need to be assessed before we can proceed. Tasks for this are listed [here](https://github.com/enram/globam-dmp/milestone/1) and include:

1. Get a file listing for `pvol` archive
2. Select subsets: first 2 days of data per radar/month for 2016, 2017, 2018 (72 days): either pvol or pre-merged scans
3. Transfer subsets to accessible FTP server
4. Merge to pvol (test 1)
5. Process with `vol2bird` to `vp` (test 2)
6. Store output `vp` files
7. Visual control of `vp` (test 3)
