# European radar data

This page describes European radar data (`pvol`) and derived data products (`vp` and `vpts`).

## License agreement {#license-agreement}

European radar data exchange is coordinated by the _European Operational Program for Exchange of Weather Radar Information ([EUMETNET/OPERA](http://eumetnet.eu/activities/observations-programme/current-activities/opera/))_. GloBAM has access to these data via the _EIG EUMETNET license agreement for Research use of OPERA data_. This license agreement grants a specific list of people (involved in ENRAM and GloBAM) access to radar data from selected countries under the following conditions (excerpt):

> The OPERA Members Data will be provided for use within the ENRAM Project for the purpose of extracting animal migration information for scientific research. The Grant of this License does not permit use of the OPERA Members Data licensed to be used for commercial purposes or exploitation for profit.

## Radars

```{r download_opera_db}
# Download OPERA database as json and store in this repo
download.file("http://eumetnet.eu/wp-content/themes/aeron-child/observations-programme/current-activities/opera/database/OPERA_Database/OPERA_RADARS_DB.json", here("data", "OPERA_RADARS_DB.json"))
```

```{r}
# Read data files
opera_radars <- fromJSON(here("data", "OPERA_RADARS_DB.json"))
opera_countries <- read_csv(here("data", "opera_countries.csv"))
```

OPERA manages a list of radar metadata (`r opera_radars %>% nrow()` radars). This **OPERA database** is available as:

* [interactive map](http://eumetnet.eu/wp-content/themes/aeron-child/observations-programme/current-activities/opera/database/OPERA_Database/index.html)
* [json file](http://eumetnet.eu/wp-content/themes/aeron-child/observations-programme/current-activities/opera/database/OPERA_Database/OPERA_RADARS_DB.json): source file for the map. This file is also [archived on GitHub](https://github.com/enram/globam-dmp/blob/master/data/OPERA_RADARS_DB.json) every time we update the DMP, so we can track changes.

The license agreement allows access to data from `r opera_countries %>% filter(in_license_agreement) %>% nrow()` countries.

```{r}
# Show table of countries
opera_radars %>%
  # Add country information
  left_join(opera_countries, by = "country") %>%
  
  # Filter on countries in license agreement
  filter(in_license_agreement) %>%
  
  # Group by country
  group_by(country, iso_code) %>%
  arrange(country) %>%
  
  # Show total number of radars + operational radars (status = 1)
  summarize(
    radars = n_distinct(number),
    operational = n_distinct(number[status == 1])
  ) %>%
  adorn_totals("row")
```

## Historical archive

GloBAM needs a **2-3 year archive** of European `pvol` and derived `vp` data to tackle its research questions. By starting from a `pvol` archive, the generation of `vp` data can be reproduced if need be, e.g. to make sure adequate vol2bird settings/versions are used.

### Source and access

- `pvol`: BALTRAD (\@ref(baltrad)) is likely to be the best source. **Restricted access**
- `vp`: ENRAM data repository (\@ref(enram-data-repo)). **Open data**
- `vpts`: ENRAM data repository (\@ref(enram-data-repo)). **Open data**

### Coverage

The BALTRAD archive needs to be assessed for coverage and quality before we can define the subset of the archive we want to use. Tasks for this are listed [here](https://github.com/enram/globam-dmp/milestone/1) and include:

1. Get a file listing for `pvol` archive
2. Select subsets: first 2 days of data per radar/month for 2016, 2017, 2018 (72 days): either pvol or pre-merged scans
3. Transfer subsets to accessible FTP server
4. Merge to pvol (test 1)
5. Process with vol2bird to `vp` (test 2)
6. Store output vp files
7. Visual control of `vp` (test 3)

## PVOL

`pvol` data are polar volume data. These are generated by weather radars and the source for vertical profiles (`vp`) of biological targets.

### Format

- Format: `hdf5` in the [OPERA ODIM_h5 format](http://www.eumetnet.eu/sites/default/files/OPERA2014_O4_ODIM_H5_v2.2.pdf).
- Granularity: a `pvol` file is typically generated every 15 minutes and contains scan data at about 20 altitudes.
- Read in bioRad with: [`read_pvolfile`](http://adokter.github.io/bioRad/reference/read_pvolfile.html).

### Source

#### BALTRAD {#baltrad}

The best source for consolidated `pvol` data is **[BALTRAD](http://baltrad.eu/)**, managed by [SMHI](https://www.smhi.se/en/research/research-departments/atmospheric-remote-sensing/contact-us-1.4801) (contact person: Günther Haase). `pvol` data are archived there, but access to that server is restricted to SMHI. For data to be available in BALTRAD, several conditions need to be met ([source](https://github.com/adokter/vol2bird/wiki/Radar-data-requirements-and-availability-for-ENRAM#future-radar-data-availability-for-enram)):

1. A country should send both reflectivity and radial velocity data to the [OPERA](http://www.eumetnet.eu/opera) data centre, which is called [ODYSSEY](http://www.eumetnet.eu/odyssey-opera-data-centre). While many countries are sending reflectivity data, radial velocities are still unavailable for many countries.
2. ODYSSEY should forward these data to the BALTRAD datahub. It is standard policy to do so, but in practice data is not yet forwarded correctly for some radars/countries. 
3. Currently BALTRAD and ODYSSEY store data at a 15 minute interval, higher resolution data is not yet available.

#### Research labs

Some research labs (e.g. UvA) also have archived `pvol` data for a subset of radars/years.

#### National weather services

More or more precise `pvol` data can be available at the national weather services as well.

### Access

**Restricted access**: access and use of `pvol` data is subject to the license agreement (\@ref(license-agreement)). These data cannot be shared further without approval from OPERA. Transferring these data to a (private) cloud platform is also sensitive.

### Coverage

Coverage of the `pvol` archive at BALTRAD is currently unknown. Availability and coverage are being gathered in [this spreadsheet](https://docs.google.com/spreadsheets/d/1dulZ4TLjxIjmUek5QP86uJ9DsZqm9d0G6vl6hnA0mw4/edit?usp=sharing).

### Up-to-dateness

`pvol` date at BALTRAD are normally up to date to within a day.

## PVOL → VP

...

## VP

`vp`

### Source {#enram-data-repo}

`vp` data generated at BALTRAD are transferred daily to the [ENRAM data repository](http://enram.github.io/data-repository/).

### Access

**Open data**: Data in ENRAM data repository are available under a [Creative Commons Zero waiver](https://creativecommons.org/publicdomain/zero/1.0/).

### Format

- Format: `hdf5` in the [ODIM bird profile format specification](https://github.com/adokter/vol2bird/wiki/ODIM-bird-profile-format-specification).
- Granularity: a `vp` file is generated for each originating `pvol` file.
- Read in bioRad with: [`read_vpfiles`](http://adokter.github.io/bioRad/reference/read_vpfiles.html).

### Coverage

The coverage of the ENRAM data repository is recorded daily in [`coverage.csv`](https://lw-enram.s3-eu-west-1.amazonaws.com/coverage.csv) and summarized here:

```{r}
vp_coverage <- read_csv("https://lw-enram.s3-eu-west-1.amazonaws.com/coverage.csv")
vp_coverage %>%
  # Create year column + rename column
  mutate(year = year(date)) %>%
  rename(radar = countryradar) %>%
  
  # Group by radar and year
  group_by(radar, year) %>%
  
  # Show number of days with vp files
  summarize(
    days = n_distinct(date)
  ) %>%
  
  # Create year columns with days values
  spread(year, days) %>%
  adorn_totals("col")
```

### Up-to-dateness

The ENRAM data repository is updated daily.

## VPTS

### Format

To be defined: https://github.com/enram/globam-dmp/issues/16

### Source

These files are not yet generated, but will be stored in the ENRAM data repository.

### Access

**Open access**

### Coverage

Not yet defined.

### Up-to-dateness

Not yet defined.
